A local-global principle for $K_2$ of $\GL_4$ and orthogonal groups

### 2016-04-19 paper.tex
Imported Andrei's source

\documentclass[12pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, enumitem, graphicx, mathtools, xfrac, tikz-cd, nameref, thmtools, hyperref}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, citestyle=numeric-comp, natbib=true, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}            
\addbibresource{paper.bib}
\usepackage[activate={true,nocompatibility}]{microtype}
\usepackage[capitalize]{cleveref}

@@

\theoremstyle{plain} \declaretheorem[name=Theorem, Refname={Theorem,Theorems}]{thm} \Crefname{thm}{Theorem}{Theorems}
\numberwithin{equation}{section}

\newtheorem{lemma}{Lemma} \numberwithin{lemma}{section} \Crefname{lemma}{Lemma}{Lemmas}
\newtheorem{cor}[lemma]{Corollary} \Crefname{cor}{Corollary}{Corollaries}
\newtheorem{prop}[lemma]{Proposition} \Crefname{prop}{Proposition}{Propositions}
\newtheorem*{thm*}{Theorem}
\newtheorem*{lemma*}{Lemma}
\theoremstyle{definition} \newtheorem{dfn}[lemma]{Definition} \Crefname{dfn}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\theoremstyle{remark} \newtheorem{rem}[lemma]{Remark} \Crefname{rem}{Remark}{Remarks}

@@

\newcommand{\catname}[1]{{\normalfont\textbf{#1}}}
\newcommand{\Ker}{\operatorname{Ker}}
\newcommand{\Cent}{\operatorname{Cent}}
\newcommand{\Img}{\operatorname{Im}}
\newcommand{\Coker}{\operatorname{Coker}}
\newcommand{\coker}{\operatorname{coker}}
\newcommand{\K}{\operatorname{\mathrm{K}}}
\newcommand{\GG}{\operatorname{\mathrm{G}}}
\newcommand{\St}{\operatorname{\mathrm{St}}}
\newcommand{\StU}{\operatorname{\hat{\mathrm{U}}}}
\newcommand{\G}{\operatorname{\mathrm{G}}}
\newcommand{\E}{\operatorname{\mathrm{E}}}
\newcommand{\M}{\operatorname{\mathrm{M}}}
\newcommand{\sr}{\operatorname{\mathrm{sr}}}
\newcommand{\Um}{\operatorname{\mathrm{Um}}}
\newcommand{\rk}{\operatorname{\mathrm{rk}}}
\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}} 
\newcommand{\rD}{\mathsf{D}} 
\newcommand{\rE}{\mathsf{E}}
\newcommand{\HH}{\mathrm{H}}
\newcommand{\eval}[4]{ev \textstyle \left[\frac{#2[#1] \rightarrow #3}{#1 \mapsto #4}\right]}
\newcommand{\ev}[3]{\eval{t}{#1}{#2}{#3}}

@@

\title{Local-global principle for $\K_2$ of Chevalley groups}
\keywords {Chevalley groups, Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09; 20G35}
\author {Andrei Lavrenov, Sergey Sinchuk}
\email {avlavrenov {\it at} gmail.com, sinchukss {\it at} gmail.com}
\date {\today}

@@

\section {Introduction}\label{intro}

\subsection{Acknowledgements}
The authors of the present paper acknowledges financial support from Russian Science Foundation grant 14-11-00297.

\section {Preliminaries}\label{prelim}

\section {Main results}

\subsection{Relative Steinberg group in rank $3$ and Tulenbaev's section map} 
The main goal of this subsection is to construct Tulenbaev's section map from~\cite[\S~2]{T} for
the linear Steinberg group of rank $3$. Combining this result with the amalgamation theorem
of the the second named author (see~\cite[Theorem~9]{S15}) we deduce centrality for the orthogonal
$\K_2$ in the next subsection.

First of all, we will need to give yet another definition of the relative linear Steinberg group 
(cf. \cite[Definitions~3.3 and 3.7]{S15}).

\begin{dfn}
 The relative Steinberg group $\St^*(n,R, I)$ is the group defined by the following two
 families generators and four families of relations.
 \begin{itemize}
  \item Generators:
  \begin{enumerate}
  \item $X^1(u, v)$, where $u \in \E(n,R) \cdot e_1$, $v\ \in I^n$ such that $v^t \cdot u = 0$;
  \item $X^2(u, v)$, where $u \in I^n$, $v \in \E(n,R) \cdot e_1$ such that $v^t \cdot u = 0$.
 \end{enumerate}
  Notice that $\phi$ maps both $X^1(u, v)$ and $X^2(u, v)$ to $T(u, v) = e_n + u \cdot v^T$.
  \item Relations:
  \begin{enumerate}
  \item $X^1(u, v) \cdot X^1(u, w) = X^1(u, v+w)$, $u \in \E(n,R) \cdot e_1$, $v, w \in I^n$;
  \item $X^2(u, v) \cdot X^2(w, v) = X^2(u+w, v)$, $u, w \in I^n$, $v \in \E(n,R) \cdot e_1$;
  \item ${}^{X^\sigma(u^2, v^2)} \! X^\tau(u^1, v^1) = X^\tau(T(u^2, v^2) \cdot u^1, T(v^2, u^2)^{-1} \cdot v^1)$, $\sigma, \tau = 1..2$;
  \item $X^1(g \cdot e_1, g^* \cdot be_2) = X^2(g \cdot be_1, g^* \cdot e_2)$ where $b\in I$ and $g^* = {g^t}^{-1}$ denotes the contragradient matrix.
 \end{enumerate}
 \end{itemize}
\end{dfn}
It is easy to check the correctness of the above definition.

\begin{lemma}
 The groups $\St^*(n, R, I)$ and $\St(n, R, I)$ are isomorphic.
\end{lemma}
\begin{proof}
 TODO: 
\end{proof}

The next step of the proof is to is construct certain elements in $\St(n, R)$ similar to Tulenbaev's elements $X_{u,v}(a)$ see~\cite[\S~1]{T}.

Let $v\in R^n$ be a column.
A column $w\in R^n$ is called \emph{$v$-decomposable} if it can be presented as a finite sum $w = \sum\limits_{i=1}^n w^i$ such that each $w^i$ has at least two zero entries. 
Denote by $D(v)$ the submodule of $R^n$ consisting of all $v$-decomposable columns.
For a column $v\in R^n$ denote by $I(v)$ the ideal of $R$ spanned by coordinates $v_1,\ldots, v_n$.

Let $u,v,w\in R^n$ be columns such that $w^tv=0$.
It is easy to check (cf.~\cite[Lemma~3.2]{Ka}) that 
$$(uv)\cdot w = \sum_{i<j}w_{ij},\ \text{where}\ w_{ij} = (w_iu_j - w_ju_i)(v_je_i - v_ie_j)\in{}\!A^n.$$
The above decomposition is called the \emph{canonical} decomposition of $(uv)\cdot w$.
In particular, the above argument shows that row $a\cdot w$ is always $v$-decomposable for $a\in I(v)$, i.e. $I(v) \cdot O(v) \subseteq D(v)$.
It is also obvious that $O(v)\subseteq O(bv)$, $D(v)\subseteq D(bv)$ for $b \in R$.
%\end{enumerate}\end{lemma}

Denote by $B^1$ the subset of $R^n \times R^n \times R$ consisting of triples $(u, v, a)$ such that $v^t \cdot u = 0$, $v \in D(u)$, $a \in I(u)$.
Denote by $B^2$ the set consisting of triples $(v, u, a)$ such that $(u, v, a) \in B^1$.

\begin{lemma} \label{lem:Zproperties}
Assume that $n \geq 4$.
One can define two families of elements $Z^\tau(u, v, a) \in \St(n, R)$, $\tau=1,2$ parametrized by $(u, v, a) \in B^\tau$ satisfying the following properties
 \begin{lemlist}
  \item \label{rel:proj} $\phi(Z^\tau(u, v, a)) = e + uav^t \in \E(n, R)$, $(u,v,a) \in B^\tau$;
  \item \label{rel:add1} $Z^{1}(u, v_1 + v_2, a) = Z^{1}(u, v_1, a) \cdot Z^{1}(u, v_2, a)$;
  \item \label{rel:add2} $Z^{2}(v_1 + v_2, u, a) = Z^{2}(v_1, u, a) \cdot Z^{2}(v_2, u, a)$;
  \item \label{rel:conj} ${}^{g}\! Z^{\tau}(u, v, a) = Z^{\tau}(\phi(g) \cdot u, \phi(g)^* \cdot v, a)$, $\tau = 1,2$, $g \in \St(n, R)$.
 \end{lemlist}
\end{lemma}
\begin{proof}
One constructs the elements $Z^1(u,v,a)$ in exactly the same way as Tulenbaev constructs his elements $X_{u,v}(a)$ (see definitions preceding~\cite[Lemma~1.2]{T}).
Next, the elements $Z^2(u,v,a)$ are constructed by ``transposing'' Tulenbaev's construction.

All the properties (with the exception of the last one in the case $n=4$) can be proved in exactly the same fashion as in~\cite[Lemma~1.3]{T}.
To check the last property in the case $n=4$ one has to use the argument of W. van~der~Kallen (cf.~\cite{Ka}).
%TODO: Write more details
\end{proof}

For the rest of this section $a$ denotes a nonnilpotent element of $R$.
\begin{lemma}
Let $u,v,w,z \in R^n$ be columns such that 
\begin{equation*} u^t \cdot v = 0,\ w^t \cdot v = c \in R,\ u^t \cdot z = 0,\ w^t \cdot z = 0. \end{equation*}
Assume, moreover, that $(u, v, a) \in B^1$, $(w, z, a) \in B^2$, then one has $$Z^1(u, ac \cdot z, a) = Z^2(ac \cdot u, z, a).$$
\end{lemma}
\begin{proof}
Direct computation using \cref{lem:Zproperties} shows that
 \begin{multline*}
 Z^2(ac \cdot u, z, a) = Z^2((e+uav^t) w, (e-vau^t)z,a) \cdot Z^2(-w,z,a) = \\
  {}^{Z^1(u,v,a)} \! Z^2(w,z,a) \cdot Z^2(-w,z,a) = [Z^1(u, v, a), Z^2(w, z, a)] = Z^1(u,v,a) \cdot {}^{Z^2(w,z,a)}\! Z^1(u,-v,a) = \\
    = Z^1(u,v,a) \cdot Z^1((e+waz^t)u,-(e-zaw^t)v,a) = Z^1(u,ac \cdot z,a). \qedhere
 \end{multline*}  
\end{proof}

\subsection{Centrality of orthogonal $K_2$.}

>>>>

\documentclass[11pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, enumitem, mathtools, xfrac, nameref, thmtools, hyperref}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber, citestyle=numeric-comp, natbib=true, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}            
\usepackage[activate={true,nocompatibility}]{microtype}
\usepackage[capitalize]{cleveref}
\usepackage[english]{babel} 
\usepackage{stmaryrd}
\usepackage[matrix,arrow,curve]{xy}

%Bibliography
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}
\newbibmacro{string+doi}[1]{\iffieldundef{doi}{\iffieldundef{url}{#1}{\href{\thefield{url}}{#1}}}{\href{http://dx.doi.org/\thefield{doi}}{#1}}}
\DeclareFieldFormat{title}{\usebibmacro{string+doi}{\mkbibemph{#1}}}
\DeclareFieldFormat[article, inproceedings, inbook, thesis]{title}{\usebibmacro{string+doi}{\mkbibquote{#1}}}
\addbibresource{paper.bib}

@@

\theoremstyle{plain} \declaretheorem[name=Theorem, Refname={Theorem,Theorems}]{tm} \Crefname{tm}{Theorem}{Theorems}
\numberwithin{equation}{section}

\newtheorem{lm}{Lemma} \numberwithin{lm}{section} \Crefname{lm}{Lemma}{Lemmas}
\newtheorem{cl}[lm]{Corollary} \Crefname{cl}{Corollary}{Corollaries}
\newtheorem{prop}[lm]{Proposition} \Crefname{prop}{Proposition}{Propositions}
\newtheorem*{tm*}{Theorem}
\newtheorem*{lm*}{Lemma}
\theoremstyle{definition} \newtheorem{df}[lm]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lm]{Example} \Crefname{example}{Example}{Examples}
\theoremstyle{remark} \newtheorem{rk}[lm]{Remark} \Crefname{rk}{Remark}{Remarks}
\newtheorem*{mtm}{Main Theorem}
\newtheorem*{atm}{Another Presentation}

@@

\newcommand{\Max}{\mathop{\mathrm{Max}}\nolimits}
\newcommand{\proj}{\mathop{\mathrm{proj}}\nolimits}
\newcommand{\Card}{\mathop{\mathrm{Card}}\nolimits}
\newcommand{\Ker}{\mathop{\mathrm{Ker}}\nolimits}
\newcommand{\Cent}{\mathop{\mathrm{Cent}}\nolimits}
\newcommand{\E}{{\mathrm{E}}}
\newcommand{\Um}{\mathop{\mathrm{Um}}\nolimits}
\newcommand{\St}{\mathop{\mathrm{St}}\nolimits}
\newcommand{\Sp}{\mathop{\mathrm{Sp}}\nolimits}
\newcommand{\Ep}{\mathop{\mathrm{Ep}}\nolimits}
\newcommand{\GL}{\mathop{\mathrm{GL}}\nolimits}
\newcommand{\SL}{\mathop{\mathrm{SL}}\nolimits}
\newcommand{\Kt}{\mathop{\mathrm{K_2}}\nolimits}
\newcommand{\Ko}{\mathop{\mathrm{K_1}}\nolimits}
\newcommand{\Ho}{\mathop{\mathrm{H_1}}\nolimits}
\newcommand{\Ht}{\mathop{\mathrm{H_2}}\nolimits}
\newcommand{\epi}{\twoheadrightarrow}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\eps}[1]{\varepsilon_{#1}}
\newcommand{\sign}[1]{\mathrm{sign}(#1)}
\newcommand{\lan}{\langle}
\newcommand{\ran}{\rangle}
\newcommand{\inv}{^{-1}}
\newcommand{\ur}[1]{\!\,^{(#1)}U_1}
\newcommand{\ps}[1]{\!\,^{(#1)}\!P_1}
\newcommand{\ls}[1]{\!\,^{(#1)}\!L_1}

\renewcommand{\labelenumi}{\theenumi{\rm)}}
\renewcommand{\theenumi}{\alph{enumi}}

@@

\title{A local-global principle for $K_2$ of $\GL_4$ and orthogonal groups}
\keywords {Steinberg groups, $K_2$-functor. {\em Mathematical Subject Classification (2010):} 19C09}

\author{Sergey Sinchuk}
\email{sinchukss {\it at} gmail.com}

\author {Andrei Lavrenov}
\email {avlavrenov {\it at} gmail.com}

\date {\today}

@@

%\begin{thebibliography}{99}
%\bibitem{Lav2} A.~Lavrenov, A local--global principle for symplectic $\mathrm K_2$, {\it to appear}.
%\bibitem{Rao} *Statement of Ravi Rao's hypothesis
%\bibitem{Swa1} *Swan's relative Steinberg group 1
%\bibitem{Swa2} *Swan's relative Steinberg group 2
%\bibitem{Swa3} *Swan's relative Steinberg group 3
%\bibitem{Tul0} *Tulenbaev's paper on almost commutative rings (cite somewhere)
%\end{thebibliography}

\section*{Inroduction}

The main goal of the present paper is to give a uniform proof of centrality of $\mathrm K_2$ for all simply-laced Chevalley groups of rank $\geq3$.\footnote{The authors acknowledge support of ER-EN-EF.}

The paper is organised as follows. In the first section we give yet another presentation for $\St(4,\,R,\,I)=\St(\mathrm A_3,\,R,\,I)$. It mimics the standart vector representation of $\SL(n,\,R)$, therefore in the first section we use representations and corresponding vector notation. Afterwards we can deduce a local--global principle for $\St(\mathrm\Phi,\,R)$ and centrality of $\Kt(\mathrm\Phi,\,R)$ for $\Phi=\mathrm A_3$. In the second section we use obtained results to prove the same theorems for any simply laced $\Phi$ of rank $\geq3$. There we switch to the root notation.

\section{Case of $\mathrm A_3$}

Tulenbaev used {\it another presentation} for a relative Steinberg group to prove a local--global principle for a linear Steinberg group~\cite{Tul}. This relative another presentation was inspired by the {\it absolute} one of van der Kallen~\cite{vdK}. However, Tulenbaev's presentation can only be used for $\St(l+1,\,R)=\St(\mathrm A_l,\,R)$, $l\geq4$.

Tulenbaev's result can be applied to the subsystems of type $\mathrm A_4$ inside $\mathrm E_l$ and ``glued'' to get a local--global principle for Steinberg groups of type $\mathrm E_l$~\cite{SCh}. To use this glueing method for groups of type $\mathrm D_l$ as well one has to work with subsystems of type $\mathrm A_3$ instead, and thus extend Tulenbaev's result to the case of $\mathrm A_3$.

Recently, the local--global principle was obtained for Steinberg groups of type $\mathrm C_l$ starting from rank 3~\cite{Lav2}. That is another motivation to consider the linear Steinberg group in rank 3.

In the present paper $R$ denotes commutative (associative, unital) ring, $l\geq3$, and $n=l+1\geq4$, $R^n$ is a free (right) $R$-module with basis $e_1,\ldots,e_n$, $\GL(n,\,R)$ is a group of invertible $n\times n$ matrices over $R$, $\E(n,\,R)$ its elementary subgroup generated by elementary transvections 
$$\{t_{ij}(r)=1+e_{ij}r\mid1\leq i\neq j\leq n,\ r\in R\},$$
where 1 stands for the identity matrix and $e_{ij}$ for the matrix unit. The Steinberg group is a group given by formal generators
$$\{x_{ij}(r)\mid1\leq i\neq j\leq n,\ r\in R\}$$
and Steinberg relations
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\begin{align}
&x_{ij}(r)x_{ij}(s)=x_{ij}(r+s),\\
&[x_{ij}(r),\,x_{hk}(s)]=1,\text{ for }h\neq j,\ k\neq i,\\
&[x_{ij}(r),\,x_{jk}(s)]=x_{ik}(rs).
\end{align}
All commutators in the present paper are left-normed, $[x,\,y]=xyx\inv y\inv$. Denote $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ a natural projection sending $x_{ij}(r)$ to $t_{ij}(r)$.

Suslin showed that (already for $n\geq3$) the elementary group coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
$$
\{t(u,\,v)=1+uv^t\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\},
$$
where $\Um(n,\,R)$ is a set of unimodular coloumns of $R^n$, i.e., such that their entries generate the unit ideal of $R$, $u^t$ stands for $u$ {\it transposed}~\cite{Sus}. This result clearly implies that $\E(n,\,R)$ is normal inside $\GL(n,\,R)$. Then, van der Kallen developed Suslin's ideas and showed that the Steinberg group $\St(n,\,R)$ is isomorphic to a group defined by a set of generators
$$
\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}
$$
and relations
\setcounter{equation}{0}
\renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w),\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v')
\end{align}
for $n\geq4$~\cite{vdK}. Then, it is also obvious that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension. Observe that van der Kallen uses pairs of a coloumn and a row $(i,\,j)$ to parametrise his generators, and we prefer two coloumns.

The local--global principle for the elementary group proven by Suslin~\cite{Sus} is a statement that an elementary matrix over a polynomial ring $g\in\E(n,\,R[X])$ becomes trivial after the evaluation ``$X=0$''
$$
g\in\Ker\big(\E(n,\,R[X])\rightarrow\E(n,\,R)\big)
$$
and after localisations
$$
g_{\mathfrak m}=1\in\E(n,\,R_{\mathfrak m}[X])
$$
for all maximal ideals $\mathfrak m$ of a base ring $R$, if and only if it is trivial from the very begining
$$
g=1\in\E(n,\,R[X]).
$$
An analogous statement about Steinberg groups was proven by Tulenbaev for $n\geq5$~\cite[Th.~2.1]{Tul}. I.e., for $n\geq5$ and $g\in\St(n,\,R[X])$ such that 
$$g(0)=1\in\St(n,\,R)\ \text{ and }\ g_{\mathfrak m}=1\in\St(n,\,R_{\mathfrak m}[X])$$
 for all $\mathfrak m\in\Max R$ holds $g=1$.

For $I\trianglelefteq R$ denote $\St(n,\,R,\,I)=\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$ {\it a relative Steinberg group}. Clearly, $\St(n,\,R[X],\,XR[X])$ is the kernel of the evaluation ``$X=0$''.

The key step in Tulenbaev's proof is constructing a homomorphism
$$
\mathrm T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X])
$$
for any (non-nilpotent) element $a\in R$, $n\geq5$ ($R_a$ stands for a principle localisation of $R$ in $a$). This $\mathrm T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated as a proof of excision property for Steinberg group in this very particular situation.

To prove the local--global principle for Steinberg groups of type $\mathrm D_l$ and centrality of $\mathrm K_2(\mathrm D_l,\,R)$ we need such a map for $n=4$ (i.e., for the group of type $\mathrm A_3$).

Tulenbaev uses the following presentaion for the relative Steinberg group, see~\cite[Prop.~1.6]{Tul}. Assume that $I\trianglelefteq R$ such that a short exact sequence
\renewcommand{\theequation}{$\dagger$}
\begin{align}
\xymatrix{
0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0
}
\end{align}
splits, i.e., $\pi\sigma=1$. The $\St(n,\,R,\,I)$ is isomorphic to a group given by generators
$$
\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}
$$
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w),\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'),\\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R).
\end{align}
Here $\Um_{n\times2}(R)$ is a set of $n\times2$ matrices $M$ such that there exists a $2\times n$ matrix $N$ with $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$. The ideal satisfying $\dagger$ is called the splitting ideal, $XR_a[X]\trianglelefteq R_a[X]$ is obviously the splitting one.

Another presentations of van der Kallen and Tulenbaev are given in terms of generators parametrised by pairs of vectors, where the first one is ``good'' (unimodular or a coloumn of an elementary matrix) and the second one is arbitrary. This allows to formulate easily an additivity by the second argument, but one has to write something similar to T3 instead of additivity by the first one.

Now, we need to construct a homomorphism
$$
\mathrm T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X]),
$$
i.e., to find elements in the target group subject to Tulenbaev's relations, but because of T3 one can do it only for $n\geq5$, see~\cite[Lemmas~1.2 and~1.3\,c)]{Tul}.

One can not parametrise generators by a pair of arbitrary vectors since it is not even known that $t(u,\,v)\in\E(n,\,R)$ for arbitrary orthogonal $u$ and $v$~\cite{Rao}.

To generalise Tulenbaev's results for $n=4$ we use a presentation with two types of generators: $F(u,\,v)$ with $u$ good and $v$ arbitrary, which are additive by the second component, and $S(u,\,v)$ with $u$ arbitrary and $v$ good, which are additive by the first component. For $u$ and $v$ both good we require that these two types of generators coincide. More precisely,

\begin{df}
for $I\trianglelefteq R$, $n\geq4$ define $\St^*(n,\,R,\,I)$ to be a group given by the set of generators
\begin{multline*}
\{F(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\,\cup\\
\cup\{S(u,\,v)\mid u\in I^n,\ v\in\E(n,\,R)e_1,\ u^tv=0\}
\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w),\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v),\\
&F(u,\,v)F(u',\,v')F(u,\,v)\inv=F(t(u,\,v)u',\,t(v,\,u)\inv v'),\\
&F(u,\,va)=S(ua,\,v)\ \forall\,a\in I,\,(u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\!\in\E(n,\,R).
\end{align}
\end{df}

One could probably wonder why we assumed only one relation on conjugated element, instead of four different relations on generators of each type conjugated by generators of each type. Three ``missing'' relations actualy hold automaticaly.

\begin{lm}
\label{allyouneedisf}
Denote $\varphi\colon\St^*(n,\,R,\,I)\rightarrow\E(n,\,R)$ the natural map, sending $F(u,\,v)\mapsto t(u,\,v)$, $S(u,\,v)\mapsto t(u,\,v)$.
\begin{enumerate}
\item 
$\St^*(n,\,R,\,I)$ is generated by the set of elements
$$
\{F(u,\,va)\mid a\in I,\ (u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\in\E(n,\,R)\};
$$
\item
for any $g\in\St^*(n,\,R,\,I)$ holds
\setcounter{equation}{2}
\renewcommand{\theequation}{R\arabic{equation}'}
\begin{align}
&gF(u,\,v)g\inv=F(\varphi(g)u,\,\varphi(g\inv)^tv);
\end{align}
\item
for any $g\in\St^*(n,\,R,\,I)$ holds
\setcounter{equation}{2}
\renewcommand{\theequation}{R\arabic{equation}''}
\begin{align}
&gS(u,\,v)g\inv=S(\varphi(g)u,\,\varphi(g\inv)^tv);
\end{align}
\item
there exists a well-defined ``transpose automorphism''
$$
F(u,\,v)^t=S(v,\,u),\quad S(u,\,v)^t=F(v,\,u).
$$
\end{enumerate}
\end{lm}

\begin{proof}
Take an arbitrary generator of the first type $F(u,\,v)$ and take $M\in\E(n,\,R)$ such that $F(u,\,v)=F(Me_1,\,M^*\tilde v)$. Here $M^*$ stands for a contragradient matrix, $M^*=(M\inv)^t$. By R1 we have
$$
F(u,\,v)=\prod\limits_{k\neq1}F(Me_1,\,M^*e_k\tilde v_k)
$$
where $\tilde v_k$ stands for $k$-th coordinate of $\tilde v=\sum e_k\tilde v_k$. Similarly, with the use of R4 we get 
$$
S(u,\,v)=\prod\limits_{k\neq1}S(Ne_k\tilde u_k,\,N^*e_1)=\prod\limits_{k\neq1}F(Ne_k,\,N^*e_1\tilde u_k)
$$
and obtain a). Obviously, b) follows from a). For c) it is sufficient to show that
$$
F(u,\,v)S(u',\,v')F(u,\,v)\inv=S(t(u,\,v)u',\,t(u,\,v)^*v').
$$
Let $S(u',\,v')=\prod F(Ne_k,\,N^*e_1\tilde u_k)$, then
\begin{multline*}
F(u,\,v)S(u',\,v')F(u,\,v)\inv=F(u,\,v)\prod F(Ne_k,\,N^*e_1\tilde u_k)F(u,\,v)\inv=\\
=\prod F(t(u,\,v)Ne_k,\,t(u,\,v)^*N^*e_1\tilde u_k)=\\
=\prod S(t(u,\,v)Ne_k\tilde u_k,\,t(u,\,v)^*N^*e_1)=S(t(u,\,v)u',\,t(u,\,v)^*v').
\end{multline*}
Finally, d) follows from c).
\end{proof}

Next, we show that for a splitting ideal $I\trianglelefteq R$ the group $\St^*(n,\,R,\,I)$ is isomorphic to $\St(n,\,R,\,I)=\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$. With this end we construct two inverse homomorphisms
$$
\xymatrix{
\St^*(n,\,R,\,I)\ar@<0.5ex>[r]^{\iota}&\St(n,\,R,\,I)\ar@<0.5ex>[l]^{\kappa}.
}
$$

Firstly, we define $\iota$ as a homomorhism to an absolute group
$$
\xymatrix{
\St^*(n,\,R,\,I)\ar@<0.0ex>[r]^{\iota}&\St(n,\,R).
}
$$
We do not need an assumption that $I$ is a splitting ideal for that.

Clearly, we map elements $F(u,\,v)$ to van der Kallen's another generators $X(u,\,v)$. To find images for $S(u,\,v)$ as well we recall that van der Kallen in~\cite[3.8--3.10]{vdK} defines also elements
$$
x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.
$$
Now, ``transpose'' the definition~\cite[3.13]{vdK}.

\begin{df}
For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$ consisiting of all elements $y\in\St(n,\,R)$ which could be decomposed as a product $\prod x(w^k,\,v)$, where $\sum w^k=u$ and $w^k=(e_pv_q-e_qv_p)c^k$ for some $c^k\in R$, $1\leq p\neq q\leq n$.
\end{df}

Since coloumns of elementary matrices are unimodular, $\overline Y(u,\,v)$ is not empty (see~\cite[3.1--3.2]{vdK}). Obviously, for $x\in\overline Y(u,\,v)$ and $y\in\overline Y(w,\,v)$ holds $xy\in\overline Y(u+w,\,v)$.

Repeating~\cite[3.14--3.15]{vdK} verbatim one shows that

\begin{lm}
for $g\in\St(n,\,R)$ holds
\begin{enumerate}
\item
$g\overline Y(u,\,v)g\inv\subseteq\overline Y(\phi(g)u,\,\phi(g)^*v)$;
\item
$\overline Y(u,\,v)$ consists of exactly one element.
\end{enumerate}
\end{lm}

The only element of $\overline Y(u,\,v)$ we denote by $Y(u,\,v)$. We map elements $S(u,\,v)\in\St^*(n,\,R,\,I)$ to $Y(u,\,v)$. To show that such a homomorphism is well-defined we check that $X(u,\,v)$ and $Y(u,\,v)$ satisfy relations R1--R4, but R1--R3 are already checked.

\begin{lm}
Consider $(u,\,v^t)=(Me_1,\,e_2^tM\inv)$, $M\in\E(n,\,R)$, $a\in I$. Then $X(u,\,va)=Y(ua,\,v)$.
\end{lm}

\begin{proof}
It suffices to show that $X(e_1,\,e_2a)=Y(e_1a,\,e_2)$. Compute the commutator $[Y(-e_3,\,e_2),\,X(e_1,\,e_3a)]$ in two ways:
$$
Y(-e_3,\,e_2)X(e_1,\,e_3a)Y(-e_3,\,e_2)\inv\cdot X(e_1,\,-e_3a)=X(e_1,\,e_2a),
$$
and
$$
Y(-e_3,\,e_2)\cdot\,X(e_1,\,e_3a)Y(e_3,\,e_2)X(e_1,\,e_3a)\inv=Y(e_1a,\,e_2).
$$
\end{proof}

Therefore, the homomorphism $\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R)$ sending $F(u,\,v)$ to $X(u,\,v)$ and $S(u,\,v)$ to $Y(u,\,v)$ is well-defined. Denote $\rho_I\colon R\epi R/I$, then one can easily see that induced morphism on Steinberg groups 
$$\rho_I^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$$
 sends $X(u,\,v)$ to $X(\rho_I(u),\,\rho_I(v))$ and $Y(u,\,v)$ to $Y(\rho_I(u),\,\rho_I(v))$. Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective so that for $x\in\E(n,\,R)e_1$ holds that $\rho_I(x)\in\E(n,\,R/I)e_1$. It follows from additivity relations that $X(u,\,0)=1=Y(0,\,v)$, thus $\mathrm{Im}(\iota)\subseteq\Ker(\rho_I^*)=\St(n,\,R,\,I)$.

We constructed $\iota$ for an arbitrary $I\trianglelefteq R$. For $I$ a splitting ideal we construct an inverse homomorphism
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I).
$$
With this end we recall a presentation of $\St(n,\,R,\,I)$ as a group with action of $\St(n,\,R)$. For a splitting ideal such a presentation was firstly obtained by Swan~\cite{Swa1,Swa2,Swa3}, and for an arbitrary $I$ by Keune~\cite{Keu} and Loday~\cite{Lod}.

For a group $G$ acting on a group $H$ from the left denote by $\!\,^gh$ the image of $h\in H$ under homomorphism corresponding by $g\in G$.
